---
title: "Run the whole pipeline with LUAD using NetNorM smoothing"
output: html_notebook
---

# Load Clinical

These script reads the clinical variables that came with the [NetNorM](https://github.com/marineLM/NetNorM) package and dichotomizes them.
```{r eval=FALSE}
rm(list = ls())
source("../R/1-createBoolClinical/1-createBoolClinical-LUAD.R")

#431 rows including header
  #20643 columns including patient ID
  clinical <- read.table(file="/Users/diamac/DDfileSystem/10-Research/Current/CLINGEN/9source/CLIGEN/testing_NetNorM/data_NetNorM/TCGA/data_preprocessed/LUAD/X_labwithclin_ns1_NANA_NA_NA.csv",
                         sep = ",", quote = "'",
                         header = TRUE, stringsAsFactors = FALSE,
                         skipNul = FALSE, blank.lines.skip = FALSE)
  
  row.names(clinical) <- clinical[,1]
  clinical <- clinical[,c(20598:20643)]
  
  
  #write.table(x = clinical[1:10,], file= "temp/10samplesClinical.csv")

  survivalVars <- clinical[,1:2]
  save(survivalVars,file= "data/survivalVars.RData")
  
  clinical <- clinical[,-c(1:2)]
  
createBoolClinical(clinical)

```
These are the list of 44 clinical variables that we will use for this analysis:

"AGE"
"SMOKING_PACK_YEARS"
"SMOKING_YEARS"
"HISTORY_OTHER_MALIGNANCY"
"RESIDUAL_TUMOR"
"LONGEST_DIMENSION"
"SHORTEST_DIMENSION"
"SPECIMEN_SECOND_LONGEST_DIMENSION"
"AJCC_NODES_PATHOLOGIC_PN"
"AJCC_TUMOR_PATHOLOGIC_PT"
"AJCC_PATHOLOGIC_TUMOR_STAGE"
"GENDER_FEMALE"
"GENDER_MALE"
"RACE_AMERICAN INDIAN OR ALASKA NATIVE"
"RACE_ASIAN"
"RACE_BLACK OR AFRICAN AMERICAN"
"RACE_WHITE"
"RACE_[Not Evaluated]"
"RACE_[Unknown]"
"TOBACCO_SMOKING_HISTORY_INDICATOR_Current Reformed Smoker"
" Duration Not Specified"
"TOBACCO_SMOKING_HISTORY_INDICATOR_Current reformed smoker for < or = 15 years"
"TOBACCO_SMOKING_HISTORY_INDICATOR_Current reformed smoker for > 15 years"
"TOBACCO_SMOKING_HISTORY_INDICATOR_Current smoker"
"TOBACCO_SMOKING_HISTORY_INDICATOR_Lifelong Non-smoker"
"ICD_O_3_HISTOLOGY_8140/3"
"ICD_O_3_HISTOLOGY_8230/3"
"ICD_O_3_HISTOLOGY_8250/3"
"ICD_O_3_HISTOLOGY_8252/3"
"ICD_O_3_HISTOLOGY_8253/3"
"ICD_O_3_HISTOLOGY_8255/3"
"ICD_O_3_HISTOLOGY_8260/3"
"ICD_O_3_HISTOLOGY_8310/3"
"ICD_O_3_HISTOLOGY_8480/3"
"ICD_O_3_HISTOLOGY_8490/3"
"ICD_O_3_HISTOLOGY_8507/3"
"ICD_O_3_HISTOLOGY_8550/3"
"ICD_O_3_SITE_C34.0"
"ICD_O_3_SITE_C34.1"
"ICD_O_3_SITE_C34.2"
"ICD_O_3_SITE_C34.3"
"ICD_O_3_SITE_C34.8"
"ICD_O_3_SITE_C34.9"
"AJCC_METASTASIS_PATHOLOGIC_PM_M0"
"AJCC_METASTASIS_PATHOLOGIC_PM_M1"


While dicotomizing, the following continuos variables where dichotomized using the jenks method:


1. SHORTEST_DIMENSION  with n=5
style: jenks
  one of 70 possible partitions of this variable into 5 classes
[0.1,0.2] (0.2,0.3] (0.3,0.4] (0.4,0.6]   (0.6,1] 
       74        85       170        78        23 

2. SPECIMEN_SECOND_LONGEST_DIMENSION
style: jenks
  one of 48,620 possible partitions of this variable into 10 classes
[0.1,0.3] (0.3,0.5] (0.5,0.6] (0.6,0.7] (0.7,0.8] (0.8,0.9] (0.9,1.1] (1.1,1.3] 
       23        54        52       122        68        29        48        13 
(1.3,1.6]   (1.6,2] 
       15         6 
       
3. LONGEST_DIMENSION
style: jenks
  one of 6,906,900 possible partitions of this variable into 10 classes
[0.1,0.4] (0.4,0.7] (0.7,0.9] (0.9,1.1] (1.1,1.3] (1.3,1.6] (1.6,1.8] (1.8,2.1] 
       14        46        64       151        44        51        22        20 
(2.1,2.6]   (2.6,3] 
       13         5 

4. SMOKING_YEARS
style: jenks
  one of 1,101,716,330 possible partitions of this variable into 10 classes
  [2,8]  (8,16] (16,22] (22,26] (26,31] (31,36] (36,42] (42,48] (48,56] (56,64] 
      4      21      16      13      26     282      39      13      12       4 

5. AGE
style: jenks
  one of 2,054,455,634 possible partitions of this variable into 10 classes
[38,46] (46,52] (52,56] (56,60] (60,63] (63,67] (67,71] (71,75] (75,80] (80,88] 
     15      24      32      63      41      54      68      65      45      23 


6. SMOKING_PACK_YEARS

style: jenks
  one of 571,350,360,240 possible partitions of this variable into 10 classes
   [0,12]   (12,23]   (23,33]   (33,44]   (44,54] (54,67.5] (67.5,84]  (84,104] 
       29        46        47       190        45        29        17        16 
(104,125] (125,154] 
        9         2


As a result we got 106 dichotomized clinical variables. Write a table explaining how.

Check Fig5, 6, 7 and 78 from "temp/10-"


```{r}
load("../data/boolClinical.RData")

```

# Load Mutation


```{r}
boolMutation <- read.table(file="/Users/diamac/DDfileSystem/10-Research/Current/CLINGEN/9source/CLIGEN/testing_NetNorM/data_NetNorM/TCGA/data_preprocessed/LUAD/X_labwithclin_ns1_NANA_NA_NA.csv",
                       sep = ",", quote = "'",
                       header = TRUE, stringsAsFactors = FALSE, 
                       skipNul = FALSE, blank.lines.skip = FALSE)

row.names(boolMutation) <- boolMutation[,1]
boolMutation <- boolMutation[,c(2:20597)]
mutationSmooth <- boolMutation

save(mutationSmooth,file= "../data/mutationSmooth.RData")

```

# Prepare data to construct tensor on Matlab

```{r}
rm(list = ls()[-c(which(ls() %in% c("mutationSmooth", "boolClinical")))])
patients <- row.names(boolClinical) 
save(patients,file= "../data/patients.RData")


tClini = t(boolClinical)
write.csv2(tClini, file = "../temp/4-Matlab-binaC.csv",row.names = FALSE)
tBina = t(mutationSmooth)
write.table(tBina, file = "../temp/10-binaM.csv", sep =';', dec = '.', row.names = FALSE )

```

# Run tensor factorization with rubik

Move the files to *"MATLAB-RUBIK/data/"* and update the readme file.

Run the script:
*"MATLAB-RUBIK/rubik/run_cligen.m"*

The output will be located on *"MATLAB-RUBIK/rubikOutput/"*


# Plot the components

```{r eval=FALSE}
source("R/3-tSNE_visualization.R'")
```

[resulting plot](https://plot.ly/~datad/7/#/)


# Clustering 

```{r}
#load data
source("R/3-writeFactors.R")
source("R/4-kmeans.R")
source("R/5-getPvalsSurvAnal.R")
```




# Other examples


```{r}
source("R/0-ConfigurationVariables.R")
minNumOfMutationsPerPatient = 5
minNumOfPatientsPerGene = 5

source("R/1-ReadRawFiles.R")
# 1- read the raw files from the 'rawData' folder and store them in 'dataFile'
readBRCAFiles (dataFile = "temp/readBRCAFiles.RData",
               rawMutationFile="rawData/data_mutations_extended.txt",
               rawClinicalFile="rawData/BRCA.clin.merged.txt")

# 2- filter the clinical variables specified in "rawData/clinical_useful.csv" and "rawData/merge_useful.csv"
filterClinicalVariables()
load("data/tensorClinical.RData")

source("R/1-createBoolMutation.R")
createBoolMutation(boolOrBin="bina")
load("data/patients.RData")
load("data/genes.RData")
load("data/binaMutation.RData")

#Filter tensor clinical
load("data/tensorClinical.RData")
row.names(tensorClinical) <- tensorClinical$patient.bcr_patient_barcode
tensorClinical <- tensorClinical[patients, ]
save(tensorClinical, file="data/tensorClinical.RData")


#Dichotomize Clinical vars


# Step 2 - Construct the tensor
source("R/2-BoolMatrices.R")
#source("R/2-Smooth_MutationM_SVD.R")
#source("R/2-Construct_tensor.R")

# Step 3 - Tensor Factorization
#source("R/3-RunTF.R")




```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
